---
title: "CMSAC Workshop"
author: "Marion Haney"
date: "2023-11-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
library(sportyR)
library(gganimate)
library(magick)
library(ggthemes)
library(gt)
```


# Load in data.

Week 1 data from the 2021 NFL season.

```{r}
games <- read_csv("2023-data/games.csv")
plays <- read_csv("2023-data/plays.csv")
players <- read_csv("2023-data/players.csv")
pffScoutingData <-read_csv("2023-data/pffScoutingData.csv")
```

Combine the tracking data for 2021 weeks 1-8.
```{r}
# path <- "2023-data/"
# tracking <- list.files(pattern = "week", path = path) %>%
#   map(read_csv) %>%
#   list_rbind()

week_files = c("2023-data/week1.csv", "2023-data/week2.csv", 
               "2023-data/week3.csv", "2023-data/week4.csv",
               "2023-data/week5.csv", "2023-data/week6.csv", 
               "2023-data/week7.csv", "2023-data/week8.csv")
tracking <- map(week_files, read_csv) %>% list_rbind()
```



# Standardize the data.

In raw football data there are 2 play directions. We want to standardize the 
data so they are moving left to right.

```{r}
tracking <- tracking |>
  mutate(
    # Make all plays go from left to right
    x = ifelse(playDirection == "left", 120 - x, x),
    y = ifelse(playDirection == "left", 160 / 3 - y, y),
    # Flip player direction and orientation
    dir = ifelse(playDirection == "left", dir + 180, dir),
    dir = ifelse(dir > 360, dir - 360, dir),
    o = ifelse(playDirection == "left", o + 180, o),
    o = ifelse(o > 360, o - 360, o)
  )
```



# Visualizing Tracking Data

Pick a play to become the case study.

Plot the football field.

```{r}
field_params <- list(field_apron = "springgreen3",
                     field_border = "springgreen3",
                     offensive_endzone = "springgreen3",
                     defensive_endzone = "springgreen3",
                     offensive_half = "springgreen3",
                     defensive_half = "springgreen3")
nfl_field <- geom_football(league = "nfl", ## this is from the sportyR package
                           display_range = "in_bounds_only",
                           x_trans = 60,
                           y_trans = 26.6667,
                           xlims = c(35, 85),
                           color_updates = field_params)
nfl_field
```


Visualizing tracking data.

FRAMES ARE IN FOLDER
```{r, message = FALSE, warning = FALSE}
# example_play <- tracking |> 
#   filter(gameId == 2021101709, playId == 1444) |> 
#   mutate(pt_color = case_when(team == "DEN" ~ "navy",
#                               team == "LV" ~ "white",
#                               team == "football" ~ "yellow"))
# nfl_field +
#   geom_point(data = example_play,
#              aes(120 - x, 160 / 3 - y),
#              size = 3,
#              color = example_play$pt_color) +
#   transition_time(example_play$frameId)
```


# Defining a metric

Performance metrics based on tracking data can be classified into two types:

1. directly-derived - directly calculated from data

2. model-based - estimated through statistical model



### Strain rate in materials science:

Ratio of stretching speed and original length. Think of a rubber band. This 
describes deformation of a material over time.

### For football pass rush:
You need a fast velocity towards the QB and close distance to QB.
This metric is directly derived.

STRAIN: -v/d --> negative velocity divided by distance

### Advantages of STRAIN: 

Directly derived measures are simpler and easier to compute. 
It is also more interpretable: "time required for a pass rusher to 
get to the QB given the current velocity and distance to the QB". 
It is also scalable, which means it can be applied to almost every passing 
plays.
This is also a continuous-time within-play metric-- meaning it can be 
calculated during the play (instantaneously) and it can take on continuous 
values.


# Implementing STRAIN

Filter data-- only ball snap and QB event. Find locations of passers and 
pass rushers. Calculate velocity and distance-- to calculate STRAIN.

Identify the starting and ending events, then collect the first observed start 
and end frames for each play.

```{r}
tracking_start_end <- tracking |>
  distinct(gameId, playId, frameId, event) |>
  mutate(is_start = as.numeric(event %in% c("autoevent_ballsnap", "ball_snap")),
  is_end = as.numeric(event %in% c("fumble", "handoff", "lateral", "autoevent_passforward",
  "pass_forward", "qb_sack", "qb_strip_sack", "run"))) |>
  group_by(gameId, playId) |>
  mutate(any_start = any(is_start == 1), any_end = any(is_end == 1)) |>
  filter(any_start, any_end) |>
  summarize(start_frame = frameId[which(is_start == 1)[1]],
  end_frame = frameId[which(is_end == 1 & frameId > start_frame)[1]]) |>
  ungroup()
```

Join with tracking data and filter frames.

```{r}
tracking <- tracking |>
  left_join(tracking_start_end, by = c("gameId", "playId")) |>
  filter(!is.na(start_frame), !is.na(end_frame),
  frameId >= start_frame, frameId <= end_frame) |>
  mutate(frameId_corrected = frameId - start_frame) 
```

Identify players and their roles. Joined the game and play metadata.

```{r}
tracking_roles <- tracking |>
  left_join(pffScoutingData, by = c("gameId", "playId", "nflId"))
```

Get passer location for every frame of the play.

```{r}
tracking_qb <- tracking_roles |>
  filter(pff_role == "Pass", pff_positionLinedUp == "QB") |>
  select(gameId, playId, frameId, x_qb = x, y_qb = y)
```

# Visualizing STRAIN


```{r}
tracking_pass_rush <- tracking_roles |>
  filter(pff_role == "Pass Rush") |>
  left_join(tracking_qb) |>
  group_by(gameId, playId, nflId) |>
  mutate(d_qb = sqrt((x - x_qb) ^ 2 + (y - y_qb) ^ 2),
         v_qb = -(d_qb - lag(d_qb)) / 0.1,
         strain = v_qb / d_qb) |>
  ungroup()
```


```{r}
example_pass_rush <- tracking_pass_rush |>
  filter(gameId == 2021101709, playId == 1444, !is.na(strain))

example_play_strain <- example_play |>
  left_join(select(example_pass_rush, gameId:frameId, strain)) |>
  mutate(
  pt_color = ifelse(nflId != 47889, pt_color, "maroon"),
  pt_color = ifelse(team != "football", pt_color, "yellow"),
  pt_size = ifelse(is.na(strain), 3, scales::rescale(strain, to = c(3, 8), from = range(example_pass_rush$strain))),
  pt_size = ifelse(team == "football", 2, pt_size)
)
```


Example play.

FRAMES ARE IN FOLDER
```{r, message = FALSE, warning = FALSE}
# nfl_field +
#   geom_point(data = example_play_strain,
#   aes(120 - x, 160 / 3 - y),
#   size = example_play_strain$pt_size,
#   color = example_play_strain$pt_color) +
#   transition_time(example_play_strain$frameId)
```


Feature curves.

```{r}
example_pass_rush |>
  filter(nflId == 47889) |>
  pivot_longer(cols = c(d_qb, v_qb, strain),
    names_to = "feature",
    values_to = "value") |>
  mutate(
    feature = fct_relevel(feature, "strain", after = Inf),
    feature = fct_recode(feature,
    "Distance" = "d_qb",
    "Velocity" = "v_qb",
    "STRAIN" = "strain")) |>
  ggplot(aes(frameId_corrected, value, color = feature)) +
  geom_line(linewidth = 1.5) +
  scale_color_manual(
  values = c("blue", "orange", "maroon")) +
  labs(x = "Number of frames since snap", 
       y = "Feature value",
       color = NULL)
```


Get an animation of both graphs: make both figures have the same duration then 
combine them together using the magick package.


# Metric Validation

How does STRAIN differ across different positions? Different play outcomes?

Which players have the best STRAIN?

Is the metric stable over time?

Does this metric provide new information?

How does STRAIN relate to previous metrics?


# STRAIN across different positions

```{r}
tracking_pass_rush |> 
  left_join(players) |>
  filter(
    frameId_corrected <= 40,
    officialPosition %in% c("OLB", "DE", "DT", "NT")
  ) |> 
  mutate(
    officialPosition = fct_relevel(officialPosition, 
                                   "OLB", "DE", "DT", "NT")
  ) |>
  group_by(frameId_corrected, officialPosition) |> 
  summarize(avg_strain = 10 * mean(strain, na.rm = TRUE)) |> 
  ggplot(aes(frameId_corrected, avg_strain, 
             color = officialPosition)) +
  geom_smooth(se = FALSE, span = 0.3, linewidth = 1) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Number of frames since snap",
       y = "Average STRAIN",
       color = "Position")
```


# STRAIN across different play outcomes

```{r}
pff_outcomes <- pffScoutingData |> 
  group_by(gameId, playId, nflId) |> 
  summarize(i_hit = sum(pff_hit, na.rm = TRUE) > 0,
            i_hurry = sum(pff_hurry, na.rm = TRUE) > 0 ,
            i_sack = sum(pff_sack, na.rm = TRUE) > 0,
            i_none = (i_hit + i_hurry + i_sack) == 0) |> 
  ungroup() |> 
  pivot_longer(i_hit:i_none,
               names_to = "outcome",
               names_prefix = "i_",
               values_to = "value") |> 
  filter(value)
```


```{r}
tracking_pass_rush |> left_join(pff_outcomes) |> 
  left_join(players) |>
  filter(frameId_corrected <= 40,
         officialPosition %in% c("OLB", "DE", "DT", "NT")
  ) |> 
  mutate(
    officialPosition = fct_relevel(officialPosition, 
                                   "OLB", "DE", "DT", "NT"),
         outcome = fct_relevel(outcome, 
                          "sack", "hit", "hurry", "none")
  ) |>
  group_by(frameId_corrected, outcome, officialPosition) |> 
  summarize(avg_strain = 10 * mean(strain, na.rm = TRUE)) |> 
  ggplot(aes(frameId_corrected, avg_strain, 
             color = officialPosition)) +
  geom_smooth(se = FALSE, span = 0.3, linewidth = 1) +
  facet_wrap(~ outcome, nrow = 1) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Number of frames since snap",
       y = "Average STRAIN",
       color = "Position") +
  theme(legend.position = "bottom")
```



# Who are the best pass rushers according to STRAIN?

```{r}
pass_rushers_eval <- pffScoutingData |>
  filter(pff_role == "Pass Rush") |>
  group_by(nflId) |>
  summarize(snaps = n_distinct(gameId, playId)) |>
  filter(snaps >= 100)
```


```{r}
ranking_pass_rush <- tracking_pass_rush |>
  filter(nflId %in% pass_rushers_eval$nflId, !is.infinite(strain)) |>
  group_by(nflId) |>
  summarize(avg_strain = 10 * mean(strain, na.rm = TRUE),
            snaps = n_distinct(gameId, playId),
            team = first(team)) |>
  left_join(players) |>
  arrange(desc(avg_strain)) |> 
  select(Player = displayName, Team = team, Position = officialPosition,
         Snaps = snaps, "Average STRAIN" = avg_strain)
```


Top 10 edge rushers:

```{r}
ranking_pass_rush |> 
  filter(Position %in% c("OLB", "DE")) |> 
  slice_head(n = 10) |> 
  gt() |> 
  fmt_number(
    columns = c("Average STRAIN"), 
    decimals = 2
  ) |>
  data_color(
    columns = c("Average STRAIN"),
    fn = scales::col_numeric(
      palette = c("lightgreen", "darkgreen"),
      domain = NULL
    )
  ) |>
  tab_header(
    title = "Top 10 Edge Rushers", 
    subtitle = "(Minimum 100 plays)"
  )
```


Top 10 interior rushers:

```{r}
ranking_pass_rush |>
  filter(Position %in% c("DT", "NT")) |>
  slice_head(n = 10) |>
  gt() |>
  fmt_number(
    columns = c("Average STRAIN"), 
    decimals = 2
  ) |>
  data_color(
    columns = c("Average STRAIN"),
    fn = scales::col_numeric(
      palette = c("lightblue", "darkblue"),
      domain = NULL
    )
  ) |> 
  tab_header(
    title = "Top 10 Interior Rushers", 
    subtitle = "(Minimum 100 plays)"
  )
```


# STRAIN and Pressure Rate

Correlate STRAIN with the pressure rate measure.
Average STRAIN vs. pressure rate: (hurries + sacks + hits)/snaps

```{r}
pff_pressure <- pffScoutingData |>
  group_by(nflId) |>
  summarize(hurries = sum(pff_hurry, na.rm = TRUE),
            sacks = sum(pff_sack, na.rm = TRUE),
            hits = sum(pff_hit, na.rm = TRUE))
```

```{r}
strain_pressure <- tracking_pass_rush |>
  filter(nflId %in% pass_rushers_eval$nflId, 
         !is.infinite(strain)) |>
  group_by(nflId) |>
  summarize(avg_strain = 10 * mean(strain, na.rm = TRUE),
            snaps = n_distinct(gameId, playId)) |>
  left_join(pff_pressure) |>
  mutate(pressure_rate = (hurries + sacks + hits) / snaps)
strain_pressure |> 
  slice_head(n = 5)
```

Find the correlation between the two metrics.

```{r}
cor(strain_pressure$avg_strain, 
    strain_pressure$pressure_rate)
```


Scatterplot of average STRAIN and pressure rate.

```{r}
strain_pressure |>
  left_join(players) |>
  mutate(
    officialPosition = fct_relevel(officialPosition, 
                                   "OLB", "DE", "DT", "NT")
  ) |>
  ggplot(aes(avg_strain, pressure_rate)) +
  geom_point(aes(color = officialPosition), size = 2) +
  geom_smooth(method = "lm", se = FALSE, 
              color = "darkblue", linewidth = 1.5) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Average STRAIN",
       y = "Pressure rate",
       color = "Position")
```



# How does STRAIN vary over time?

Get the average STRAIN for the first 4 weeks and the last 4 weeks-- see if they 
are correlated with each other.

```{r}
strain_weeks <- tracking_pass_rush |>
  filter(nflId %in% pass_rushers_eval$nflId, 
         !is.infinite(strain)) |>
  left_join(games) |>
  mutate(i_week = ifelse(week > 4, "last4", "first4")) |>
  group_by(nflId, i_week) |>
  summarize(avg_strain = 10 * mean(strain, na.rm = TRUE)) |>
  ungroup() |>
  pivot_wider(names_from = i_week, values_from = avg_strain)
strain_weeks |> slice_head(n = 5)
```


```{r}
cor(strain_weeks$first4, strain_weeks$last4)
```


Visualize average STRAIN over time.

```{r}
strain_weeks |>
  left_join(players) |>
  mutate(
    officialPosition = fct_relevel(officialPosition, 
                                   "OLB", "DE", "DT", "NT")
  ) |>
  ggplot(aes(first4, last4)) +
  geom_point(aes(color = officialPosition), size = 2) +
  geom_smooth(method = "lm", se = FALSE, 
              color = "darkblue", linewidth = 1.5) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Average STRAIN (first 4 weeks)",
       y = "Average STRAIN (last 4 weeks)",
       color = "Position")
```


# Is STRAIN more predictive of pressure than 'pressure' itself?

Yes, based on this study's findings.


# Multilevel Model: Specification

Study resampled data using bootstrap over team drives within games. Used this 
to model STRAIN performance of players with confidence intervals/distributions.

TJ Watt is pretty good! Yay go Steelers.



