---
title: "Edge Intensity Rating"
author: "Shane Hauck, Marion Haney, Devin Basley, Vinay Maruri"
date: "2024-1-5"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gganimate)
theme_set(theme_bw())
```

```{r}
#read in data
all_data <- read_csv("data/all_processed_data.csv")
```



# Filter out penaltys

```{r}
# Remove plays where playDescription includes "PENALTY"
all_data <- all_data %>% filter(!grepl("PENALTY", playDescription))
```


# Filter out unusually big changes in direction
```{r}
all_data <- all_data %>% filter(total_change_at_max_change < 30)
```

# Filter out unusual EPA
```{r}
all_data <- all_data %>% filter(abs(expectedPointsAdded) <= 6) %>% filter(!is.na(expectedPointsAdded))
```


# Get information about the ball carrier during time of interest
```{r}
ballcarrier <- all_data %>%
  group_by(gamePlayId, frameId) %>%
  filter(frameId > start_frameId & frameId < end_frameId) %>%
  filter(nflId == ballCarrierId) %>%
  mutate(
    ballCarrier_x = x,
    ballCarrier_y = y,
    ballCarrier_s = s,
    ballCarrier_a = a,
    ballCarrier_o = o
  ) %>%
  select(
    gamePlayId, frameId, ballCarrier_x, ballCarrier_y,
    ballCarrier_s, ballCarrier_a, ballCarrier_o
  )

all_data <- full_join(all_data, ballcarrier, by = c("gamePlayId", "frameId"))
```

# Get information about players being considered as edge defenders
```{r}
edgedefenders <- all_data %>%
  group_by(gamePlayId, frameId) %>%
  filter(
    club != possessionTeam,
    club != "football"
  ) %>%
  # Filter players with a y coordinate prior to snap not relative to play side
  mutate(
    y_at_5 = ifelse(frameId == 5, y, NA)
  ) %>%
  group_by(nflId, gamePlayId) %>%
  fill(y_at_5, .direction = c("updown")) %>%
  ungroup() %>%
  filter(ifelse(playSide == "left", y_at_5 > ballY_preSnap, y_at_5 < ballY_preSnap)) %>%
  # Calculate player distance from ball carrier
  mutate(distance_from_ballCarrier = sqrt((x - ballCarrier_x)^2 + (y - ballCarrier_y)^2)) %>%
  # Filter players so they are on the playside edge of the ball carrier
  filter(ifelse(playSide == "left", y >= ballCarrier_y, y <= ballCarrier_y)) %>%
  # Filter players so they are within 5 yards of ball carrier or is closest player
  #group_by(gamePlayId, frameId) %>%
  #filter(distance_from_ballCarrier <= 5 |
  #         distance_from_ballCarrier == min(distance_from_ballCarrier, na.rm = T)) %>%
  #ungroup() %>%
  #mutate(edgeSetterId = nflId) %>%
  mutate(
    # Calculate angle to ball carrier
    angle_toballCarrier = abs(atan2(y - ballCarrier_y, x - ballCarrier_x) * (180 / pi)),
    # Calculate and adjust orientation relative to the ball carrier
    relative_o_to_ballCarrier = (((o - (atan2(y - ballCarrier_y, x - ballCarrier_x) *
      (180 / pi))) %% 360) +
      ifelse(playSide == "left", 180, 0)) %% 360
  ) %>%
  select(
    gamePlayId, frameId, playSide, nflId, distance_from_ballCarrier,
    x, y, angle_toballCarrier, s, a, relative_o_to_ballCarrier
  )

all_data <- full_join(all_data, edgedefenders %>% select(-c(x,y,s,a, playSide)), by = c("gamePlayId", "frameId", "nflId"))
```

# Information about ball carriers and edge defenders

```{r}
ballcarrier_edgedefenders <- all_data %>%
  select(ballCarrierId, colnames(ballcarrier), colnames(edgedefenders), total_change, start_frameId, end_frameId, frameId_at_max_change) %>%
  filter(frameId >= start_frameId & frameId <= frameId_at_max_change + 1) %>%
  select(-start_frameId, -end_frameId, -frameId_at_max_change) %>%
  mutate(
    # Calculate edge defender's angle relative to ball carrier's ideal path
    angle_toIdealPath = ifelse(playSide == "left",
      (360 - ((angle_toballCarrier - ballCarrier_o) + 360) %% 360),
      ((angle_toballCarrier - ballCarrier_o) + 360) %% 360
    )
  ) %>%
  mutate(
    # Convert orientations to radians
    ballCarrier_o_rad = ballCarrier_o * (pi / 180),
    edgeDef_o_rad = relative_o_to_ballCarrier * (pi / 180),

    # Calculate components of velocity
    ballCarrier_velocity_x = ballCarrier_s * cos(ballCarrier_o_rad),
    ballCarrier_velocity_y = ballCarrier_s * sin(ballCarrier_o_rad),
    edgeDef_velocity_x = s * cos(edgeDef_o_rad),
    edgeDef_velocity_y = s * sin(edgeDef_o_rad),

    # Calculate vector from edge setter to ball carrier
    vector_to_ballCarrier_x = ballCarrier_x - x,
    vector_to_ballCarrier_y = ballCarrier_y - y,

    # Project speed of edge setter towards ball carrier
    edgeDef_speed_towards_ballCarrier = (edgeDef_velocity_x * vector_to_ballCarrier_x +
      edgeDef_velocity_y * vector_to_ballCarrier_y) /
      sqrt(vector_to_ballCarrier_x^2 + vector_to_ballCarrier_y^2),
    
    # Calculate components of acceleration
    ballCarrier_acceleration_x = ballCarrier_a * cos(ballCarrier_o_rad),
    ballCarrier_acceleration_y = ballCarrier_a * sin(ballCarrier_o_rad),
    edgeDef_acceleration_x = a * cos(edgeDef_o_rad),
    edgeDef_acceleration_y = a * sin(edgeDef_o_rad),

    # Calculate vector from edge defender to ball carrier
    vector_to_ballCarrier_x = ballCarrier_x - x,
    vector_to_ballCarrier_y = ballCarrier_y - y,

    # Project acceleration of edge defender towards ball carrier
    edgeDef_acceleration_towards_ballCarrier = (edgeDef_acceleration_x * vector_to_ballCarrier_x +
                                                edgeDef_acceleration_y * vector_to_ballCarrier_y) /
                                               sqrt(vector_to_ballCarrier_x^2 + vector_to_ballCarrier_y^2),

  ) %>%
  select(
    gamePlayId, frameId, ballCarrierId, nflId,
    # defender stats
    distance_from_ballCarrier, relative_o_to_ballCarrier,
    angle_toIdealPath, edgeDef_speed_towards_ballCarrier, edgeDef_acceleration_towards_ballCarrier,
    # ball carrier stats
    total_change, ballCarrier_s, ballCarrier_a
  ) %>%
  na.omit()
```

# Create Edge Intensity Rating

```{r}
EIR <- ballcarrier_edgedefenders %>%
  mutate(
    # Normalization of variables on a scale of 0 to 1 (0 worse, 1 better)
    
    # Scaling variables where lower is better
    
    # Distance to ball carrier
    scale_dist = 1 - 
      (distance_from_ballCarrier - min(distance_from_ballCarrier, na.rm = TRUE)) / 
      (max(distance_from_ballCarrier, na.rm = TRUE) - min(distance_from_ballCarrier, na.rm = TRUE)),
    
    # Speed of defender
    scale_s = 1 - 
      (ballCarrier_s - min(ballCarrier_s, na.rm = TRUE)) / 
      (max(ballCarrier_s, na.rm = TRUE) - min(ballCarrier_s, na.rm = TRUE)),
    
    # Acceleration of defender
    scale_a = 1 - 
      (ballCarrier_a - min(ballCarrier_a, na.rm = TRUE)) / 
      (max(ballCarrier_a, na.rm = TRUE) - min(ballCarrier_a, na.rm = TRUE)),

    # Scaling variables where higher is better
    
    # Speed of defender towards ball carrier relative to how close a player is to ball carrier
    edgeDef_speed_towards_bc_relative = edgeDef_speed_towards_ballCarrier * scale_dist,
    scale_speed_towards_bc = 
      ((edgeDef_speed_towards_bc_relative - min(edgeDef_speed_towards_bc_relative, na.rm = TRUE)) /
      (max(edgeDef_speed_towards_bc_relative, na.rm = TRUE) - min(edgeDef_speed_towards_bc_relative, na.rm = TRUE))),
    
    # Acceleration of defender towards ball carrier
    edgeDef_acceleration_towards_bc_relative = edgeDef_acceleration_towards_ballCarrier * scale_dist,
    scale_acceleration_towards_bc = 
      ((edgeDef_acceleration_towards_bc_relative - min(edgeDef_acceleration_towards_bc_relative, na.rm = TRUE)) /
      (max(edgeDef_acceleration_towards_bc_relative, na.rm = TRUE) - min(edgeDef_acceleration_towards_bc_relative, na.rm = TRUE))),
    
    # Total change of ball carrier
    scale_total_change = 
      (total_change - min(total_change, na.rm = TRUE)) / 
      (max(total_change, na.rm = TRUE) - min(total_change, na.rm = TRUE)),
    
    # Scaling degrees variables
    
    # Orientation of defender relative to ball carrier
    scale_o_to_bc = ((cos(pi * relative_o_to_ballCarrier / 180) + 1) / 2) * scale_dist,  
    
    # Angle of defender relative to ball carrier's ideal path
    scale_angle_to_path = ((cos(pi * angle_toIdealPath / 180) + 1) / 2) * scale_dist,
    
    
  ) %>%
  # Calculate the mean of the scaled values
  reframe(gamePlayId, frameId, nflId,
    edge_intensity_rating = rowMeans(select(., starts_with("scale_")), na.rm = TRUE),
    # Normalize EIR to a scale of 0 to 100
    edge_intensity_rating = ((edge_intensity_rating - min(edge_intensity_rating, na.rm = TRUE)) / 
      (max(edge_intensity_rating, na.rm = TRUE) - min(edge_intensity_rating, na.rm = TRUE))) * 100
  ) 

all_data <- full_join(all_data, EIR, by = c("gamePlayId", "frameId", "nflId"))
```


# Create variable for highest EIR at each frame

```{r}
all_data <- all_data %>%
  group_by(gamePlayId, frameId) %>%
  mutate(
    max_EIR_atFrame = max(edge_intensity_rating, na.rm = TRUE),
    max_EIR_atFrame = ifelse(max_EIR_atFrame == -Inf, NA, max_EIR_atFrame),
    frames_from_start = frameId - start_frameId
  ) %>%
  mutate(
    edgeSetterId_atFrame = ifelse(edge_intensity_rating == max_EIR_atFrame, nflId, NA)
  ) %>%
  fill(edgeSetterId_atFrame, .direction = "updown") %>%
  ungroup() 
```

# Create varibles for EIR at max total change of direction

```{r}
all_data <- all_data %>%
  group_by(gamePlayId) %>%
  mutate(
    max_EIR_atMaxChange = max_EIR_atFrame[which.max(total_change)],
    edgeSetterId_atMaxChange = edgeSetterId_atFrame[which.max(total_change)],
    frames_from_start_atMaxChange = frames_from_start[which.max(total_change)]
  ) %>%
  ungroup() 
```

```{r}
maxChangesEIR <- all_data %>% 
  filter(frameId == frameId_at_max_change) %>%
  distinct(gamePlayId, playSide, ballCarrierId, edgeSetterId_atMaxChange, 
           frames_from_start_atMaxChange, total_change_at_max_change, relative_total_change,
           max_EIR_atMaxChange, expectedPointsAdded)
``` 





# EDA

```{r}
ggplot(maxChangesEIR, aes(x = max_EIR_atMaxChange, y = total_change_at_max_change)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  labs(
    title = "Edge Intensity Rating vs. Total Change of Direction",
    y = "Total Change of Direction",
    x = "Edge Intensity Rating"
  )
```

```{r}
ggplot(maxChangesEIR, aes(x = max_EIR_atMaxChange, y = expectedPointsAdded)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  labs(
    title = "Edge Intensity Rating vs. Expected Points Added",
    y = "Expected Points Added",
    x = "Edge Intensity Rating"
  )
```

```{r}
EIR_TC_plot <- maxChangesEIR %>%
  select(total_change_at_max_change, max_EIR_atMaxChange) %>%
  mutate(bucket = cut(total_change_at_max_change, breaks = seq(0, 30, 3))
         ) %>% 
  group_by(bucket) %>% 
  mutate(bucket_angle_min = as.numeric(sub("\\((.*),.*", "\\1", as.character(bucket)))) %>%
  mutate(meanEIR = mean(max_EIR_atMaxChange))

ggplot(EIR_TC_plot, aes(x = bucket_angle_min, y = 1, fill = meanEIR)) +
  geom_col() +
  coord_polar() +
  scale_x_continuous(limits = c(0,180)) +
  scale_y_continuous(limits = c(0,1)) +
  scale_fill_gradient2(high = "darkblue", low = "firebrick4", midpoint = mean(EIR_TC_plot$meanEIR, na.rm = TRUE)) +
  labs(title = "Edge Intensity Rating by Greatest Total Change of Direction", 
       subtitle = "For frames where the ball carrier had the highest change in total direction", 
       x = "Total Change", y = "", 
       fill = "EIR") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.title.x = element_text(size = 14),
        plot.title = element_text(size = 16),
        plot.subtitle = element_text(size = 12),
        legend.key.size = unit(1, "cm"))
```

```{r}
EIR_RTC_plot <- maxChangesEIR %>%
  select(relative_total_change, max_EIR_atMaxChange) %>%
  mutate(bucket = cut(relative_total_change, breaks = seq(-30, 30, 3))
         ) %>% 
  group_by(bucket) %>% 
  mutate(bucket_angle_min = as.numeric(sub("\\((.*),.*", "\\1", as.character(bucket)))) %>%
  mutate(meanEIR = mean(max_EIR_atMaxChange))

ggplot(EIR_RTC_plot, aes(x = bucket_angle_min, y = 1, fill = meanEIR)) +
  geom_col() +
  coord_polar(start = 66) +
  scale_x_continuous(limits = c(-90,90)) +
  scale_y_continuous(limits = c(0,1)) +
  scale_fill_gradient2(high = "darkblue", low = "firebrick4", midpoint = mean(EIR_RTC_plot$meanEIR, na.rm = TRUE)) +
  labs(title = "Edge Intensity Rating by Greatest Relative Total Change", 
       x = "Relative Total Change", y = " ", 
       subtitle = "Values < 0 are Cutbacks, Values > 0 are Bounces",
       fill = "EIR")  +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.title.x = element_text(size = 14),
        plot.title = element_text(size = 16),
        plot.subtitle = element_text(size = 12),
        legend.key.size = unit(1, "cm"))
```


