---
title: "Tables"
author: "Shane Hauck, Marion Haney"
date: "2024-01-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

```{r}
BC_ES_data <- read_csv("data/BC_and_ES_information.csv") %>% select(-1)
players <- read_csv("data/players.csv")

tables_df <- left_join(BC_ES_data, players %>% select(nflId, position, displayName), by = c("ballCarrierId" = "nflId")) %>%
  rename(BC_position = position, BC_name = displayName) %>%
  left_join(players %>% select(nflId, position, displayName), by = c("edgeSetterId_atMaxChange" = "nflId")) %>%
  rename(ES_position = position, ES_name = displayName) 
```

# Making tables for the Shiny app Defense view

We want to show top edge setters for a team based on the EIR metric.

Table needs:
playSide, defensiveTeam, ES_name, setting_freq, meanTime, meanEIR, meanEPA

```{r}
# Make a time variable: 1 frame = 10ms, 10 frames = 1 sec
eir_df <- tables_df %>%
  mutate(time = frames_from_start_atMaxChange / 10)

# How often the player is setting the edge for their team
eir_df <- eir_df %>%
  group_by(defensiveTeam, playSide) %>%
  mutate(n_plays = n()) %>%
  group_by(defensiveTeam, playSide, ES_name) %>%
  mutate(n_player_plays = n()) %>%
  group_by(ES_name) %>%
  mutate(setting_freq = n_player_plays / n_plays)

# Average EIR and EPA for each player when they are setting the edge
eir_df <- eir_df %>%
  group_by(playSide, ES_name) %>%
  mutate(meanEIR = mean(max_EIR_atMaxChange), 
         meanEPA = mean(expectedPointsAdded),
         meanTime = mean(time))

# Select only the values we want for the table
eir_df <- eir_df %>%
  select(defensiveTeam, ES_name, playSide, n_plays, n_player_plays,
         setting_freq, meanTime, meanEIR, meanEPA) %>%
  unique() %>%
  arrange(defensiveTeam, playSide, ES_name)
```

Save df as eir_epa.csv
```{r}
write.csv(eir_df, "edge-setting-epa/eir_epa.csv")
```

Checking EIR values:
```{r}
hist(eir_df$meanEIR)
```


# Making tables for the Shiny app Offense view

## Team Offense View

We want to show the mean EPA and frequency for the team's performance on 
inside zone run plays, specifically bounces, dives, and cutbacks.

We define bounces, dives, and cutbacks using relative_total_change:

RIGHT SIDE
Cutback: < -30, -10)
Dive: -10, 10)
Bounce: 10, > 30

LEFT SIDE
Bounce: < -30, -10)
Dive: -10, 10)
Cutback: 10, > 30

Table data needed: team, playSide, type, frequency, meanEPA

```{r}
play_types_R <- c("Cutback", "Dive", "Bounce")
play_types_L <- c("Bounce", "Dive", "Cutback")
ranges <- c(-45, -10, 10, 40)
```

First, add play type based on relative_total_change ranges.

```{r}
team_table_R <- tables_df %>%
  filter(playSide == "right") %>%
  # Make a column for play type: type
  mutate(type = cut(relative_total_change, breaks = ranges, 
                    labels = play_types_R, include.lowest = TRUE))

team_table_L <- tables_df %>%
  filter(playSide == "left") %>%
  # Make a column for play type: type
  mutate(type = cut(relative_total_change, breaks = ranges, 
                    labels = play_types_L, include.lowest = TRUE))

team_table_df <- rbind(team_table_R, team_table_L)
```

```{r}
team_table_df <- team_table_df %>%
  # We want frequency at the play side level: for left and right
  group_by(possessionTeam, playSide) %>%
  mutate(n_plays = n()) %>%
  group_by(possessionTeam, playSide, type) %>%
  mutate(n_plays_type = n()) %>%
  # Make column for play type frequency: freq
  mutate(freq = n_plays_type / n_plays)

team_table_df <- team_table_df %>%
  # Make column for mean EPA for each team's play type on left and right
  group_by(possessionTeam, playSide, type) %>%
  mutate(meanEPA = mean(expectedPointsAdded)) %>%
  # Select only columns we need
  select(possessionTeam, playSide, type, n_plays, n_plays_type, freq, meanEPA)

team_table_df <- unique(team_table_df) %>% arrange(possessionTeam, playSide, type)
```

Save df as team_routes_epa.csv
```{r}
write.csv(team_table_df, "edge-setting-epa/team_routes_epa.csv")
```


<!-- ## Player Offense View -->

<!-- We want to show "players to watch", i.e. the offense team's runningbacks  -->
<!-- and how effective they are for different routes on inside zone run plays. -->

<!-- Table data needed: team, BC_name, playSide, type, meanEPA, most_common_route -->

<!-- ```{r} -->
<!-- player_table_df <- tables_df %>% -->
<!--   # Make a column for play type: type -->
<!--   mutate(type = cut(relative_total_change, breaks = ranges,  -->
<!--                     labels = play_types, include.lowest = TRUE)) %>% -->
<!--   group_by(BC_name, playSide) %>% -->
<!--   # Make a column for player play type frequency for each side -->
<!--   mutate(n_plays = n()) %>% -->
<!--   group_by(BC_name, playSide, type) %>% -->
<!--   mutate(n_plays_type = n()) %>% -->
<!--   mutate(freq = n_plays_type / n_plays) -->

<!-- player_table_df <- player_table_df %>% -->
<!--   # Make column for mean EPA for each team's play type on left and right -->
<!--   group_by(BC_name, type, playSide) %>% -->
<!--   mutate(meanEPA = mean(expectedPointsAdded)) %>% -->
<!--   # Select only columns we need -->
<!--   select(possessionTeam, BC_name, BC_position, playSide, type,  -->
<!--          n_plays, n_plays_type, freq, meanEPA) -->

<!-- player_table_df <- unique(player_table_df) %>% arrange(possessionTeam, BC_name,  -->
<!--                                                        playSide, type) -->
<!-- ``` -->

<!-- Save df as player_routes_epa.csv -->
<!-- ```{r} -->
<!-- write.csv(player_table_df, "edge-setting-epa/player_routes_epa.csv") -->
<!-- ``` -->

