---
title: "Triangles"
author: "Shane Hauck, Marion Haney, Devin Basley, Vinay Maruri"
date: "2023-12-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(progress)
library(tidyverse)
library(gganimate)
theme_set(theme_bw())

library(sportyR)
field_params <- list(
  field_apron = "springgreen3",
  field_border = "springgreen3",
  offensive_endzone = "springgreen3",
  defensive_endzone = "springgreen3",
  offensive_half = "springgreen3",
  defensive_half = "springgreen3"
)

nfl_field <- function(xlims) {
  geom_football(
    league = "nfl",
    display_range = "full",
    x_trans = 60,
    y_trans = 26.6667,
    xlims = xlims,
    color_updates = field_params
  )
}
```


# Reading in the cleaned data: just inside zone run plays
## USE WK2_SameDirection_TrackData for analysis of week 2

```{r}
wk2_track_data <- read.csv("data/WK2_SameDirection_TrackData.csv")
```

Make gamePlayId which is a combination of gameId and playId. This helps us 
identify each unique play.
```{r}
wk2_track_data$gamePlayId <- paste(wk2_track_data$gameId, wk2_track_data$playId)
```

# FINDING POINTS OF THE TRIANGLE

We want to identify 3 points of the triangle which represents the ballCarrier's
path: start, end, and furthest.

Save these coordinates in the data. Then, we can calculate the angle of change.

## Start
(X,Y) of ballCarrier when event == “handoff”

start_x, start_y, start_frameId

```{r}
start <- wk2_track_data %>%
  filter(nflId == ballCarrierId) %>%
  group_by(gamePlayId) %>%
  filter(event == "handoff") %>%
  reframe(
    start_x = x,
    start_y = y,
    start_frameId = frameId,
    gamePlayId = gamePlayId
  )

wk2_track_data <- merge(wk2_track_data, start, by = "gamePlayId")
```

## End
If: event includes “tackle”, (X,Y) of ballCarrier when event == “tackle”
Else: location where ballCarrier breaks gap point
Shane’s soccer code: checks every player pairing and sees if ballCarrier line 
crosses in-between lines

Do when the player breaks gap point first, because they could be tackled 
after that. Then, if end_x and end_y are NA for that play, check those plays 
for the tackle events.

end_x, end_y, end_frameId


```{r}
# Breaks the gap point
# Create Variables to get ball Carriers X and Y coordinates
wk2_track_data <- wk2_track_data %>%
  group_by(gamePlayId) %>%
  mutate(
    ballCarrierX = ifelse(nflId == ballCarrierId, x, NA),
    ballCarrierY = ifelse(nflId == ballCarrierId, y, NA)
  ) %>%
  fill(ballCarrierX, .direction = c("updown")) %>%
  fill(ballCarrierY, .direction = c("updown"))


# Function to check intersection between two line segments
# Returns a list with the intersection status and point coordinates
is_line_segment_intersect <- function(x1, y1, x2, y2, x3, y3, x4, y4) {
  # Handle cases of vertical line segments
  vertical1 <- x1 == x2
  vertical2 <- x3 == x4
  # [Additional logic for vertical lines can be added here]

  # For non-vertical lines, calculate intersection point
  if (!vertical1 && !vertical2) {
    slope1 <- (y2 - y1) / (x2 - x1)
    intercept1 <- y1 - slope1 * x1
    slope2 <- (y4 - y3) / (x4 - x3)
    intercept2 <- y3 - slope2 * x3

    # Check if non-parallel lines intersect
    if (slope1 != slope2) {
      x_intersect <- (intercept2 - intercept1) / (slope1 - slope2)
      y_intersect <- slope1 * x_intersect + intercept1

      # Ensure intersection point lies within both line segments
      if (x_intersect >= min(x1, x2) && x_intersect <= max(x1, x2) &&
        x_intersect >= min(x3, x4) && x_intersect <= max(x3, x4)) {
        return(list(intersect = TRUE, x = x_intersect, y = y_intersect))
      }
    }
  }

  # No intersection or calculation not possible
  return(list(intersect = FALSE, x = NA, y = NA))
}

# Function to determine if ball carrier splits the gap between players
# Returns intersection status, point, and frameId
check_intersection <- function(frameId, ballCarrierStartX, ballCarrierStartY, ballCarrierEndX, ballCarrierEndY, players_df) {
  # Rank players by y-coordinate for adjacency check
  players_df$rank <- rank(players_df$y)
  player_pairs <- combn(nrow(players_df), 2, simplify = FALSE)

  # Iterate through player pairs for intersection check
  for (pair in player_pairs) {
    player1 <- players_df[pair[1], ]
    player2 <- players_df[pair[2], ]

    # Check adjacency based on y-coordinate rank
    if (abs(player1$rank - player2$rank) == 1) {
      result <- is_line_segment_intersect(
        ballCarrierStartX, ballCarrierStartY, ballCarrierEndX, ballCarrierEndY,
        player1$x, player1$y, player2$x, player2$y
      )
      if (result$intersect) {
        # Return intersection details including frameId
        return(c(result, frameId = frameId))
      }
    }
  }

  # No intersection found among adjacent players
  return(list(intersect = FALSE, x = NA, y = NA, frameId = frameId))
}

# Calculate total number of groups for progress bar
total_groups <- wk2_track_data %>%
  filter(position %in% c("T", "G", "C") | nflId == ballCarrierId) %>%
  count(gamePlayId, frameId) %>%
  ungroup() %>%
  summarise(total = sum(n)) %>%
  pull(total)

# Initialize progress bar for process monitoring
pb <- progress_bar$new(total = total_groups, format = " Processing [:bar] :percent :elapsed")

# Process data to find first gap split by ball carrier
splitGap <- wk2_track_data %>%
  filter(position %in% c("T", "G", "C") | nflId == ballCarrierId) %>%
  group_by(gamePlayId) %>%
  arrange(frameId) %>%
  do({
    pb$tick() # Update progress bar
    with(., {
      ball_carrier_positions <- filter(., nflId == ballCarrierId)
      if (nrow(ball_carrier_positions) > 2) {
        for (i in 2:nrow(ball_carrier_positions)) {
          start_pos <- ball_carrier_positions[i - 1, ]
          end_pos <- ball_carrier_positions[i, ]
          current_frameId <- end_pos$frameId
          other_players <- filter(., nflId != ballCarrierId & frameId == current_frameId)

          # Check for intersection and return result with frameId
          if (nrow(other_players) > 1) {
            result <- check_intersection(current_frameId, start_pos$x, start_pos$y, end_pos$x, end_pos$y, other_players)
            if (result$intersect) {
              return(data.frame(split_gap = TRUE, x_intersect = result$x, y_intersect = result$y, frameId = result$frameId))
            }
          }
        }
        return(data.frame(split_gap = FALSE, x_intersect = NA, y_intersect = NA, frameId = NA))
      } else {
        return(data.frame(split_gap = FALSE, x_intersect = NA, y_intersect = NA, frameId = NA))
      }
    })
  }) %>%
  ungroup() %>%
  # Summarize results across gamePlayIds
  group_by(gamePlayId) %>%
  summarise(split_gap = any(split_gap), x_intersect = first(x_intersect), y_intersect = first(y_intersect), frameId_intersect = first(frameId), .groups = "drop")
```

```{r}
wk2_track_data <- merge(wk2_track_data, splitGap, by = "gamePlayId")
```


```{r}
# Just the first of tackle, first_contact, touchdown, or out of bounds that occurs
triangle_end_point <- wk2_track_data %>%
  filter(nflId == ballCarrierId) %>%
  group_by(gamePlayId) %>%
  filter(event %in% c("tackle", "first_contact", "touchdown", "out_of_bounds") |
    frameId == frameId_intersect) %>%
  filter(frameId == min(frameId)) %>%
  reframe(
    end_x = x,
    end_y = y,
    end_frameId = frameId,
    gamePlayId = gamePlayId
  )
```


```{r}
# Merge back to the data
wk2_track_data <- merge(wk2_track_data, triangle_end_point, by = "gamePlayId")
```


## Furthest
If ball carrier didn't split a gap do between start_frameID and frame before first contact. 
If ball carrier split a gap do between start_frameID and frameId_intersect.
Them take max of the absolute value of:
1. Start Y - max Y
2. Start Y - min Y

furthest_x, furthest_y, furthest_frameId

```{r}
# Figure out furthest y coordinate (max of either direction)
furthest_y <- wk2_track_data %>%
  group_by(gamePlayId) %>%
  filter(nflId == ballCarrierId) %>%
  filter(frameId > start_frameId & frameId < end_frameId) %>%
  reframe(
    max_y = max(y),
    min_y = min(y),
    start_y = start_y,
    abs_max_y = abs(start_y - max_y),
    abs_min_y = abs(start_y - min_y),
    furthest_y = ifelse(abs_max_y >= abs_min_y, max_y, min_y),
    x = x,
    y = y,
    frameId = frameId
  )
```

```{r}
# Grab the x coordinate and frame id where we see furthest_y
furthest <- furthest_y %>%
  group_by(gamePlayId) %>%
  filter(y == furthest_y) %>%
  reframe(
    furthest_x = x,
    furthest_y = y,
    furthest_frameId = frameId,
    gamePlayId = gamePlayId
  ) %>%
  group_by(gamePlayId) %>%
  filter(furthest_frameId == min(furthest_frameId))
```

```{r}
# Merge back to the data
wk2_track_data <- merge(wk2_track_data, furthest, by = "gamePlayId")
```


# Find angle of change
The angle we want is the inside angle at the furthest point.
If we have vertices F, S, and E (furthest, start, end) we want angle F, 
which is made from line segments EF and FS.

1. angle FSE = arctan((end_y - furthest_y)/ (end_x - furthest_y))
2. angle F = 180 - angle FSE

Radians to degree conversion: degree = radians * (180/pi)

```{r}
# Function to calculate the angle of change, angle F
find.angle.f <- function(x_s, y_s, x_e, y_e, x_f, y_f) {
  # Convert input values to numeric
  x_s <- as.numeric(x_s)
  y_s <- as.numeric(y_s)
  x_e <- as.numeric(x_e)
  y_e <- as.numeric(y_e)
  x_f <- as.numeric(x_f)
  y_f <- as.numeric(y_f)

  # Calculate the angle FSE in radians
  fse_rad <- atan2(y_e - y_f, x_e - x_f)

  # Convert the angle from radians to degrees
  fse_deg <- fse_rad * (180 / pi)

  # Calculate the angle at vertex F (angle EF and FS)
  f_deg <- 180 - fse_deg

  return(f_deg)
}
```

```{r}
# Apply the function that calculates the angle to the data
wk2_track_data <- wk2_track_data %>%
  rowwise() %>%
  mutate(
    angle_furthest = find.angle.f(start_x, start_y, end_x, end_y, furthest_x, furthest_y),
    angle_furthest = ifelse(playSide == "left", 360 - angle_furthest, angle_furthest)
  )
```

```{r}
ggplot(wk2_track_data %>% distinct(gamePlayId, angle_furthest, playSide), aes(x = angle_furthest)) +
  geom_histogram() +
  facet_wrap(~playSide)
```

```{r}
ggplot(wk2_track_data %>% distinct(gamePlayId, angle_furthest, playSide, expectedPointsAdded), aes(x = angle_furthest, y = expectedPointsAdded)) +
  geom_point() +
  facet_wrap(~playSide) +
  geom_smooth()
```


# Finding the greatest change in direction

```{r}
# For each frame, calculate mean coordinates of surrounding frames and use them to calculate angles
ball_carrier_grouped <- wk2_track_data %>%
  filter(nflId == ballCarrierId) %>%
  group_by(gamePlayId) %>%
  arrange(gamePlayId, frameId) %>%
  mutate(
    mean_cur_x = rowMeans(cbind(lag(x, 1, default = NA), x, lead(x, 1, default = NA)), na.rm = TRUE),
    mean_cur_y = rowMeans(cbind(lag(y, 1, default = NA), y, lead(y, 1, default = NA)), na.rm = TRUE),
    mean_prev_x = rowMeans(cbind(lag(x, 2, default = NA), lag(x, 3, default = NA), lag(x, 4, default = NA)), na.rm = TRUE),
    mean_prev_y = rowMeans(cbind(lag(y, 2, default = NA), lag(y, 3, default = NA), lag(y, 4, default = NA)), na.rm = TRUE),
    mean_next_x = rowMeans(cbind(lead(x, 2, default = NA), lead(x, 3, default = NA), lead(x, 4, default = NA)), na.rm = TRUE),
    mean_next_y = rowMeans(cbind(lead(y, 2, default = NA), lead(y, 3, default = NA), lead(y, 4, default = NA)), na.rm = TRUE)
  ) %>%
  ungroup()

# Calculate the angle for each frame using mean coordinates
ball_carrier_grouped <- ball_carrier_grouped %>%
  rowwise() %>%
  mutate(
    angle = find.angle.f(mean_prev_x, mean_prev_y, mean_next_x, mean_next_y, mean_cur_x, mean_cur_y),
    angle = ifelse(playSide == "left", 360 - angle, angle),
    angle_diff_from_180 = abs(180 - angle)
  ) %>%
  ungroup()

# Find the frame with the greatest change in direction and the corresponding angle
max_angle_change <- ball_carrier_grouped %>%
  group_by(gamePlayId) %>%
  filter(frameId >= start_frameId & frameId <= end_frameId) %>%
  filter(angle_diff_from_180 == max(angle_diff_from_180, na.rm = TRUE)) %>%
  reframe(playSide,
    angle_at_max_diff = first(angle),
    frameId_at_max_diff = first(frameId)
  ) %>%
  distinct(gamePlayId, playSide, angle_at_max_diff, frameId_at_max_diff)
```

```{r}
wk2_track_data <- merge(wk2_track_data, max_angle_change %>% select(-playSide), by = "gamePlayId")
```


```{r}
ggplot(max_angle_change, aes(x = angle_at_max_diff)) +
  geom_histogram() +
  facet_wrap(~playSide)

ggplot(wk2_track_data %>% distinct(gamePlayId, angle_at_max_diff, playSide, expectedPointsAdded), aes(x = angle_at_max_diff, y = expectedPointsAdded)) +
  geom_point() +
  facet_wrap(~playSide) +
  geom_smooth()

ggplot(wk2_track_data %>% distinct(gamePlayId, start_frameId, frameId_at_max_diff, playSide, expectedPointsAdded), aes(x = frameId_at_max_diff - start_frameId, y = expectedPointsAdded)) +
  geom_point() +
  facet_wrap(~playSide) +
  geom_smooth()
```


```{r}
wk2_track_data %>%
  group_by(gamePlayId) %>%
  reframe(frame_comparison = abs(frameId_at_max_diff - furthest_frameId)) %>%
  distinct(gamePlayId, frame_comparison) %>%
  ggplot(aes(x = frame_comparison)) +
  geom_histogram()
```

# Where the greates change in direction occurs calculate that angle with the end point

```{r}
full_angle <- wk2_track_data %>%
  filter(nflId == ballCarrierId) %>%
  filter(frameId == frameId_at_max_diff) %>%
  mutate(
    angle_greatestChange = find.angle.f(start_x, start_y, end_x, end_y, x, y),
    angle_greatestChange = ifelse(playSide == "left", 360 - angle_greatestChange, angle_greatestChange)
  )
```

```{r}
ggplot(full_angle, aes(x = angle_greatestChange)) +
  geom_histogram() +
  facet_wrap(~playSide) +
  geom_vline(xintercept = 180, linetype = "dashed")

ggplot(full_angle, aes(x = angle_at_max_diff, y = expectedPointsAdded)) +
  geom_point() +
  facet_wrap(~playSide) +
  geom_smooth() +
  geom_vline(xintercept = 180, linetype = "dashed")

ggplot(full_angle %>% filter(expectedPointsAdded != min(expectedPointsAdded)), aes(x = frameId_at_max_diff - start_frameId)) +
  geom_histogram() +
  facet_wrap(~playSide)
```

```{r}
wk2_track_data <- merge(wk2_track_data, full_angle %>% select(gamePlayId, angle_greatestChange), by = "gamePlayId")
```






```{r}
# Function to calculate the smallest angular difference
smallest_angle_diff <- function(angle1, angle2) {
  diff <- abs(angle1 - angle2)
  ifelse(diff > 180, 360 - diff, diff)
}


ballcarrier <- wk2_track_data %>%
  group_by(gamePlayId, frameId) %>%
  filter(frameId > start_frameId & frameId < end_frameId) %>%
  filter(nflId == ballCarrierId) %>%
  mutate(
    ballCarrier_x = x,
    ballCarrier_y = y,
    ballCarrier_s = s,
    ballCarrier_a = a,
    ballCarrier_o = o
  ) %>%
  mutate(
    # Standardize orientation based on playSide
    ballCarrier_standardized_orientation = ifelse(playSide == "left", 360 - ballCarrier_o, ballCarrier_o)
  ) %>%
  group_by(gamePlayId) %>%
  mutate(
    # Calculate previous frame's standardized orientation
    prev_standardized_orientation = lag(ballCarrier_standardized_orientation, 1, default = first(ballCarrier_standardized_orientation)),
    # Calculate the smallest difference in orientation between frames
    ballCarrier_orientation_change = smallest_angle_diff(ballCarrier_standardized_orientation, prev_standardized_orientation)
  ) %>%
  select(
    gamePlayId, frameId, ballCarrier_x, ballCarrier_y,
    ballCarrier_s, ballCarrier_a, ballCarrier_o, ballCarrier_orientation_change, ballCarrier_standardized_orientation
  )

wk2_track_data <- full_join(wk2_track_data, ballcarrier, by = c("gamePlayId", "frameId"))

edgedefenders <- wk2_track_data %>%
  group_by(gamePlayId, frameId) %>%
  filter(
    club != possessionTeam,
    club != "football"
  ) %>%
  # calculate player distance from ball carrier
  mutate(distance_from_ballCarrier = sqrt((x - ballCarrier_x)^2 + (y - ballCarrier_y)^2)) %>%
  filter(ifelse(playSide == "left", y >= ballCarrier_y, y <= ballCarrier_y)) %>%
  filter(distance_from_ballCarrier <= 15) %>%
  #mutate(edgeSetterId = nflId) %>%
  mutate(
    angle_toballCarrier = abs(atan2(y - ballCarrier_y, x - ballCarrier_x) * (180 / pi)),
    # Calculate and adjust orientation relative to the ball carrier
    relative_o_to_ballCarrier = (((o - (atan2(y - ballCarrier_y, x - ballCarrier_x) *
      (180 / pi))) %% 360) +
      ifelse(playSide == "left", 180, 0)) %% 360
  ) %>%
  select(
    gamePlayId, frameId, playSide, nflId, distance_from_ballCarrier,
    x, y, angle_toballCarrier, s, a, relative_o_to_ballCarrier
  )



wk2_track_data <- full_join(wk2_track_data, edgedefenders %>% select(-c(x,y,s,a, playSide)), by = c("gamePlayId", "frameId", "nflId"))

ballcarrier_edgedefenders <- wk2_track_data %>%
  select(ballCarrierId, colnames(ballcarrier), colnames(edgedefenders), start_frameId, end_frameId) %>%
  left_join(ball_carrier_grouped %>% select(gamePlayId, frameId, angle_diff_from_180), by = c("gamePlayId", "frameId")) %>%
  filter(frameId >= start_frameId & frameId <= end_frameId) %>%
  select(-start_frameId, -end_frameId) %>%
  mutate(
    # Calculate edge defender's angle relative to ball carrier's ideal path
    angle_toIdealPath = ifelse(playSide == "left",
      (360 - ((angle_toballCarrier - ballCarrier_o) + 360) %% 360),
      ((angle_toballCarrier - ballCarrier_o) + 360) %% 360
    )
  ) %>%
  mutate(
    # Convert orientations to radians
    ballCarrier_o_rad = ballCarrier_o * (pi / 180),
    edgeDef_o_rad = relative_o_to_ballCarrier * (pi / 180),

    # Calculate components of velocity
    ballCarrier_velocity_x = ballCarrier_s * cos(ballCarrier_o_rad),
    ballCarrier_velocity_y = ballCarrier_s * sin(ballCarrier_o_rad),
    edgeDef_velocity_x = s * cos(edgeDef_o_rad),
    edgeDef_velocity_y = s * sin(edgeDef_o_rad),

    # Calculate vector from edge setter to ball carrier
    vector_to_ballCarrier_x = ballCarrier_x - x,
    vector_to_ballCarrier_y = ballCarrier_y - y,

    # Project speed of edge setter towards ball carrier
    edgeDef_speed_towards_ballCarrier = (edgeDef_velocity_x * vector_to_ballCarrier_x +
      edgeDef_velocity_y * vector_to_ballCarrier_y) /
      sqrt(vector_to_ballCarrier_x^2 + vector_to_ballCarrier_y^2),

    # Calculate vector in the direction of ball carrier's movement
    ballCarrier_movement_vector_x = ballCarrier_velocity_x - edgeDef_velocity_x,
    ballCarrier_movement_vector_y = ballCarrier_velocity_y - edgeDef_velocity_y,

    # Normalize the ball carrier's movement vector to create a unit vector
    magnitude_ballCarrier_movement = sqrt(ballCarrier_movement_vector_x^2 + ballCarrier_movement_vector_y^2),
    ballCarrier_movement_unit_x = ballCarrier_movement_vector_x / magnitude_ballCarrier_movement,
    ballCarrier_movement_unit_y = ballCarrier_movement_vector_y / magnitude_ballCarrier_movement,

    # Project edge setter's velocity onto ball carrier's movement direction
    edgeDef_speed_towards_ballCarrierPath = (edgeDef_velocity_x * ballCarrier_movement_unit_x +
      edgeDef_velocity_y * ballCarrier_movement_unit_y)
  ) %>%
  select(
    gamePlayId, frameId, ballCarrierId, ballCarrier_s, ballCarrier_a, angle_diff_from_180, ballCarrier_orientation_change,
    nflId,
    distance_from_ballCarrier, angle_toballCarrier, relative_o_to_ballCarrier,
    angle_toIdealPath, edgeDef_speed_towards_ballCarrier, edgeDef_speed_towards_ballCarrierPath
  ) %>%
  na.omit()
```

# Create Edge Intensity Rating

```{r}
EIR <- ballcarrier_edgedefenders %>%
  mutate(
    # Scaling variables where lower is better
    scale_s = 1 - (ballCarrier_s - min(ballCarrier_s, na.rm = TRUE)) / (max(ballCarrier_s, na.rm = TRUE) - min(ballCarrier_s, na.rm = TRUE)),
    scale_a = 1 - (ballCarrier_a - min(ballCarrier_a, na.rm = TRUE)) / (max(ballCarrier_a, na.rm = TRUE) - min(ballCarrier_a, na.rm = TRUE)),
    scale_dist = 1 - (distance_from_ballCarrier - min(distance_from_ballCarrier, na.rm = TRUE)) / (max(distance_from_ballCarrier, na.rm = TRUE) - min(distance_from_ballCarrier, na.rm = TRUE)),
    scale_angle = 1 - (angle_toballCarrier - min(angle_toballCarrier, na.rm = TRUE)) / (max(angle_toballCarrier, na.rm = TRUE) - min(angle_toballCarrier, na.rm = TRUE)),

    # Scaling variables where higher is better
    scale_angle_diff = (angle_diff_from_180 - min(angle_diff_from_180, na.rm = TRUE)) / (max(angle_diff_from_180, na.rm = TRUE) - min(angle_diff_from_180, na.rm = TRUE)),
    scale_orientation_change = (ballCarrier_orientation_change - min(ballCarrier_orientation_change, na.rm = TRUE)) / (max(ballCarrier_orientation_change, na.rm = TRUE) - min(ballCarrier_orientation_change, na.rm = TRUE)),
    scale_speed_towards_bc = (edgeDef_speed_towards_ballCarrier - min(edgeDef_speed_towards_ballCarrier, na.rm = TRUE)) / (max(edgeDef_speed_towards_ballCarrier, na.rm = TRUE) - min(edgeDef_speed_towards_ballCarrier, na.rm = TRUE)),
    scale_speed_towards_path = (edgeDef_speed_towards_ballCarrierPath - min(edgeDef_speed_towards_ballCarrierPath, na.rm = TRUE)) / (max(edgeDef_speed_towards_ballCarrierPath, na.rm = TRUE) - min(edgeDef_speed_towards_ballCarrierPath, na.rm = TRUE)),

    # Scaling degrees variables
    scale_o_to_bc = (cos(pi * relative_o_to_ballCarrier / 180) + 1) / 2,  # Normalized to range from 0 to 1
    scale_angle_to_path = (cos(pi * angle_toIdealPath / 180) + 1) / 2,
    
    # Include distance twice so it has doubled the weight
    scale_dist_doubled = scale_dist
  ) %>%
  # Calculate the mean of the scaled values
  reframe(gamePlayId, frameId, nflId,
    edge_intensity_percentage = rowMeans(select(., starts_with("scale_")), na.rm = TRUE)
  ) %>%
  mutate(
    edge_intensity_rating = ((edge_intensity_percentage - min(edge_intensity_percentage)) / (max(edge_intensity_percentage) - min(edge_intensity_percentage))) * 100  
  )
```

```{r}
ggplot(EIR,
       aes(x = edge_intensity_percentage)) +
  geom_histogram() +
  labs(title = "Distribution of Edge Intensity Percentage", x = "Edge Intensity Percentage", y = "Count")

ggplot(EIR, aes(x = edge_intensity_rating)) +
  geom_histogram() +
  labs(title = "Distribution of Edge Intensity Ratings", x = "Edge Intensity Rating", y = "Count")


```


# Save the data
Saving the data as wk2_triangles.csv

```{r}
# Save dataframe as CSV
write.csv(wk2_track_data, file = "data/wk2_triangles.csv", row.names = FALSE)
```
